\rhead{ML401: Genetic Algorithms}
\phantomsection 
\section*{Genetic Algorithms (ML401)}
\label{major:ML_4}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 3 (2-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Foundations of Evolutionary Computation:} Biological evolution principles and Darwinian natural selection, Genetic algorithms as population-based stochastic optimization, Search space representation and fitness landscapes, Schema theorem and building block hypothesis, Convergence properties and implicit parallelism, Comparison with gradient-based optimization methods, No Free Lunch theorem implications for EAs.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Genetic Algorithm Components and Operators:} Encoding schemes (binary, real-valued, permutation, tree-based), Fitness function design and scaling (rank, linear, exponential), Selection methods (roulette wheel, tournament, rank, elitism), Crossover operators (single-point, uniform, arithmetic, blend), Mutation strategies (bit-flip, Gaussian, swap, scramble), Reproduction and generational replacement policies.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Advanced Genetic Algorithm Techniques:} Adaptive parameter control (self-adaptive mutation rates, dynamic crossover probabilities), Niching methods (fitness sharing, crowding, island models), Multi-objective optimization (Pareto dominance, NSGA-II, SPEA2), Constraint handling (penalty functions, repair algorithms, decoder-based), Hybridization with local search (memetic algorithms), Parallel GA architectures.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Genetic Programming and Symbolic Regression:} Tree-based representation for GP, Primitive sets and strongly-typed GP, Crossover and mutation for tree structures (subtree exchange, hoist), Bloat control mechanisms (Pareto front approximation, operator equal fitness), Automatic programming and symbolic regression, Koza's GP parameters and ramped half-and-half initialization, Multi-expression programming (MEP).

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Real-World Applications and Engineering:} Combinatorial optimization (traveling salesman problem, scheduling, knapsack), Function optimization (Rastrigin, Ackley, multimodal deceptive functions), Machine learning applications (feature selection, neural architecture search, hyperparameter optimization), Engineering design optimization (truss structures, aerodynamic shapes), Parameter tuning for deep learning and robotics controllers.

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: Basic GA Implementation} \\
Focuses on implementing single-objective GA for function optimization, experimenting with encoding schemes and selection methods.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Selection and Crossover Analysis} \\
Focuses on comparative study of selection operators and crossover types, measuring convergence speed and solution quality.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Fitness Landscape Visualization} \\
Focuses on visualizing fitness landscapes, analyzing local/global optima, and deceptive problem characteristics.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Multi-objective NSGA-II Implementation} \\
Focuses on Pareto front approximation, diversity preservation, and non-dominated sorting for engineering trade-offs.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Genetic Programming Basics} \\
Focuses on tree-based GP for symbolic regression, analyzing expression trees and bloat control effectiveness.

\vspace{0.2cm}

\noindent\textbf{Lab 6: TSP Optimization} \\
Focuses on permutation encoding for traveling salesman problem, comparing GA with nearest neighbor heuristics.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Job Shop Scheduling} \\
Focuses on constraint handling and schedule optimization using priority-based genetic representations.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Neural Network Hyperparameter Tuning} \\
Focuses on using GA for architecture search and hyperparameter optimization of MLPs.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Memetic Algorithm Hybridization} \\
Focuses on combining GA with gradient descent/local search for improved convergence.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Real-World Engineering Optimization} \\
Focuses on complete GA/GP pipeline for structural/mechanical design optimization problems.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
