\rhead{RB401: Robot Operating System (ROS)}
\phantomsection 
\section*{Robot Operating System (ROS) (RB401)}
\label{major:RB_4}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 3 (2-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{ROS Fundamentals and Communication:} ROS design principles (modularity, distributed systems, hardware abstraction), ROS master (roscore) and graph concepts, Node architecture and lifecycle, ROS topics (publish/subscribe model, QoS policies), ROS services (request/response), ROS parameters (global configuration), ROS messages and serialization (msg/srv/action definitions), ROS command line tools (rostopic, rosservice, rosparam).

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{ROS Packages, Workspaces, and Tools:} Catkin build system and workspace structure (src, devel, build), Package manifest (package.xml), CMakeLists.txt configuration, ROS launch files (parameter substitution, remapping, namespaces), RViz visualization (displays, panels, plugins), rqt tools suite (rqt\_graph, rqt\_console, rqt\_plot), Gazebo simulator integration and URDF modeling.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{TF and Robot Kinematics:} TF (Transform) library and frame transforms, Static/dynamic broadcasters and listeners, TF trees and visualization, Forward kinematics using KDL, Denavit-Hartenberg parameters, Inverse kinematics solvers, ROS Control framework (hardware\_interface, controller\_manager), Joint trajectory controllers.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{ROS Navigation Stack:} ROS navigation architecture (move\_base, costmaps, global/local planners), AMCL (Adaptive Monte Carlo Localization), SLAM (gmapping, cartographer, hdl\_graph\_slam), Path planning algorithms (A*, DWA, TEB), Recovery behaviors and costmap configuration, Nav2 stack improvements (Behavior Trees, SMAC planner), Map server and lifecycle management.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{ROS 2, Perception, and Advanced Topics:} ROS 2 DDS middleware (CycloneDDS, Fast-RTPS), ROS 2 nodes and executables (colcon build system), Composition and launch improvements, ROS-Industrial (URDF for manufacturing), Perception pipelines (PCL point clouds, OpenCV integration), MoveIt motion planning (OMPL planners, collision checking), ROS realtime extensions and micro-ROS for embedded systems.

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: ROS Workspace and Basic Nodes} \\
Focuses on creating catkin workspaces, publisher/subscriber nodes, launch files, and parameter servers.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Custom Message Types and Services} \\
Focuses on defining msg/srv/action files, implementing service clients/servers, and message serialization.

\vspace{0.2cm}

\noindent\textbf{Lab 3: TF Transform Broadcasting} \\
Focuses on static/dynamic TF broadcasters, frame hierarchies, and RViz visualization of transform trees.

\vspace{0.2cm}

\noindent\textbf{Lab 4: URDF Robot Modeling} \\
Focuses on creating URDF models with links/joints, meshes, sensors, and Gazebo simulation plugins.

\vspace{0.2cm}

\noindent\textbf{Lab 5: ROS Control Implementation} \\
Focuses on hardware interfaces, joint controllers, and gazebo\_ros plugins for simulated robot control.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Navigation Stack Configuration} \\
Focuses on map creation, AMCL localization, path planning, and obstacle avoidance tuning.

\vspace{0.2cm}

\noindent\textbf{Lab 7: SLAM Implementation} \\
Focuses on gmapping/cartographer SLAM with real/simulated robot data, loop closure analysis.

\vspace{0.2cm}

\noindent\textbf{Lab 8: MoveIt Motion Planning} \\
Focuses on MoveIt setup assistant, planning scenes, collision objects, and pick/place pipelines.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Perception Pipeline Development} \\
Focuses on point cloud processing, object detection, and camera-LiDAR fusion.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Complete ROS Robotics Application} \\
Focuses on autonomous navigation project integrating localization, mapping, planning, and manipulation.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
