\rhead{SE401: Software Testing \& Quality Assurance}
\phantomsection 
\section*{Software Testing \& Quality Assurance (SE401)}
\label{major:SE_4}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 3 (2-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Testing Fundamentals and Test Planning:} Testing principles (early testing, defect clustering, pesticide paradox, context-dependent testing), Test process integration with SDLC, Test planning components (scope, strategy, resources, schedule, risks), Test design techniques classification (black-box, white-box, experience-based), IEEE 829 test documentation standards, Test maturity models and TMMi levels.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Black-Box Testing Techniques:} Equivalence partitioning and boundary value analysis, Decision table testing and state transition testing, Use case testing and exploratory testing, Classification tree method, Pairwise testing and combinatorial test design, Domain testing and risk-based testing strategies, Test oracle problem and pairwise compatibility testing.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{White-Box Testing and Code Coverage:} Statement, branch/decision, and condition coverage metrics, Multiple condition coverage (MC/DC), Path testing and basis path testing, Loop testing strategies, Data flow testing (define-use paths), Mutation testing fundamentals and mutant operators, Code coverage tools and interpretation of coverage reports.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Test Automation and Frameworks:} Test automation pyramid (unit, integration, UI), Automation frameworks (data-driven, keyword-driven, hybrid), Page Object Model and Screen Object Model, Selenium WebDriver architecture and locators, API testing with REST Assured/Postman, Performance testing basics (load, stress, endurance), Continuous testing in CI/CD pipelines.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Quality Assurance and Test Management:} Software quality models (ISO 9126, ISO 25010), Quality gates and defect leakage metrics, Test management tools (JIRA, TestRail, Zephyr), Defect lifecycle and severity/priority classification, Test estimation techniques (3-point estimation, Work Breakdown Structure), Shift-left testing and DevOps test automation, Testability design principles and observability.

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: Test Case Design Workshop} \\
Focuses on applying equivalence partitioning, boundary value analysis, and decision tables to create comprehensive test cases from requirements.

\vspace{0.2cm}

\noindent\textbf{Lab 2: State Transition Testing} \\
Focuses on modeling system behavior as state machines and deriving test cases from state transition diagrams.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Code Coverage Analysis} \\
Focuses on instrumenting code with coverage tools, achieving target coverage levels, and interpreting coverage reports.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Mutation Testing Experiment} \\
Focuses on generating mutants, measuring mutation score, and analyzing test suite effectiveness against mutants.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Selenium Web Automation} \\
Focuses on Page Object Model implementation, cross-browser testing, and handling dynamic web elements.

\vspace{0.2cm}

\noindent\textbf{Lab 6: API Testing Automation} \\
Focuses on REST API testing with data-driven frameworks, authentication testing, and contract testing.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Performance Test Planning} \\
Focuses on creating load test scenarios, ramp-up strategies, and analyzing performance bottlenecks.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Test Management Workflow} \\
Focuses on defect lifecycle management, test execution tracking, and generating test metrics reports.

\vspace{0.2cm}

\noindent\textbf{Lab 9: CI/CD Test Integration} \\
Focuses on integrating automated tests into Jenkins/GitHub Actions pipelines with parallel execution.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Complete Test Automation Project} \\
Focuses on end-to-end test automation suite development following test pyramid principles with quality gates.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
