\rhead{RB501: Communication \& IoT}
\phantomsection 
\section*{Communication \& IoT (RB501)}
\label{major:RB_5}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 3 (2-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{IoT Communication Fundamentals:} IoT reference model (devices, gateway, edge, cloud), Wireless communication principles (modulation, spectrum, bandwidth), Short-range protocols (Bluetooth LE, Zigbee, Z-Wave), LPWAN technologies (LoRaWAN, NB-IoT, LTE-M, Sigfox), Frequency bands and regulatory domains, Power consumption models and duty cycle constraints, Network topologies (star, mesh, tree).

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{IoT Protocols and Messaging:} MQTT protocol (QoS levels, retained messages, last will), CoAP for constrained devices (DTLS security, observe pattern), HTTP/2 and WebSockets for IoT, AMQP and STOMP messaging, DDS for robotics real-time communication, Protocol benchmarking (overhead, latency, throughput), Header compression techniques (6LoWPAN).

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{IoT Security and Device Management:} Device authentication (X.509 certificates, pre-shared keys, token-based), Secure communication (TLS/DTLS, mbedTLS), Firmware Over-The-Air (OTA) updates, Device lifecycle management (provisioning, decommissioning), Secure boot and attestation, Key management in constrained environments, Zero-touch provisioning (ZTP).

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Edge Computing and Gateway Architectures:} Edge vs. cloud computing trade-offs, Gateway patterns (protocol translation, data aggregation, local processing), EdgeX Foundry and Eclipse IoT frameworks, Time-Sensitive Networking (TSN) for industrial IoT, Containerized edge computing (K3s, microk8s), Data preprocessing at edge (filtering, transformation, analytics).

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{IoT Platforms and Robotics Integration:} AWS IoT Core (Greengrass, Device Shadow), Azure IoT Hub (Device Twin, DPS), Google Cloud IoT (Cloud IoT Edge), ROS2 DDS integration with IoT protocols, Multi-protocol gateways for robotics fleets, Digital twins and simulation-in-the-loop, Industrial IoT protocols (OPC UA, Modbus TCP, PROFINET).

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: Bluetooth LE and Zigbee Interfacing} \\
Focuses on BLE GATT services, Zigbee mesh networking, and multi-protocol sensor node development.

\vspace{0.2cm}

\noindent\textbf{Lab 2: LoRaWAN Network Deployment} \\
Focuses on LoRa gateway setup, end-device registration, and long-range sensor data transmission.

\vspace{0.2cm}

\noindent\textbf{Lab 3: MQTT/CoAP Client Implementation} \\
Focuses on QoS testing, secure connections, and protocol interoperability benchmarking.

\vspace{0.2cm}

\noindent\textbf{Lab 4: IoT Security Implementation} \\
Focuses on certificate-based authentication, secure OTA updates, and device attestation.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Edge Gateway Development} \\
Focuses on protocol translation, local data processing, and cloud-edge synchronization.

\vspace{0.2cm}

\noindent\textbf{Lab 6: ROS2-IoT Integration} \\
Focuses on DDS bridge to MQTT, multi-robot fleet communication, and cloud data streaming.

\vspace{0.2cm}

\noindent\textbf{Lab 7: AWS/Azure IoT Platform Integration} \\
Focuses on device shadows, rule engine configuration, and serverless IoT backends.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Industrial Protocol Implementation} \\
Focuses on OPC UA client/server, Modbus gateways, and PROFINET device simulation.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Digital Twin Development} \\
Focuses on ROS-Gazebo simulation synchronization with physical IoT devices.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Complete IoT Robotics Solution} \\
Focuses on multi-device IoT robotics deployment with edge-cloud orchestration.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
