\rhead{CX101: Intro to Programming Lab}
\phantomsection 
\section*{Introduction to Programming (CS101)}
\label{course:CX101}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 2 (0-0-2)

\subsection*{Course Content}
\noindent\textbf{Lab 1: Basic I/O and Arithmetic} \
Focuses on the implementation of basic Input/Output operations and arithmetic logic to construct a Mini Calculator, establishing familiarity with the programming environment. \\

\vspace{0.2cm}

\noindent\textbf{Lab 2: Control Flow Patterns} \
Focuses on the application of conditional statements and looping constructs to generate complex visual patterns (Diamond, Pascal’s Triangle) and process iterative data reports. \\

\vspace{0.2cm}

\noindent\textbf{Lab 3: Pointers and Memory Management} \
Focuses on direct memory manipulation using pointers, implementing dynamic allocation for structures, and performing efficient matrix operations. \\

\vspace{0.2cm}

\noindent\textbf{Lab 4: Recursion and Combinatorics} \
Focuses on the implementation of recursive algorithms to solve complex combinatorial problems (Subsets, Combination Sum) and array partitioning strategies. \\

\vspace{0.2cm}

\noindent\textbf{Lab 5: Classes and Encapsulation} \
Focuses on Object-Oriented fundamentals, implementing Classes and Objects to demonstrate data hiding, Constructor logic, and the Deep Copy mechanism. \\

\vspace{0.2cm}

\noindent\textbf{Lab 6: Inheritance and Polymorphism} \
Focuses on hierarchical class structures, implementing Inheritance for code reusability and Virtual Functions to achieve Runtime Polymorphism. \\

\vspace{0.2cm}

\noindent\textbf{Lab 7: Advanced Class Features and Exceptions} \
Focuses on robustness and design patterns, implementing Static members, the Singleton Pattern, and Exception Handling for error management. \\

\vspace{0.2cm}

\noindent\textbf{Lab 8: The Standard Template Library (STL)} \
Focuses on leveraging pre-built generic containers (Maps, Stacks, Queues) and standard algorithms to efficiently process and organize data. \\

\vspace{0.2cm}

\noindent\textbf{Lab 9: Modern C++ Features} \
Focuses on contemporary C++ practices, specifically implementing automatic memory management with Smart Pointers and using Lambda expressions for concise logic. \\

\vspace{0.2cm}

\noindent\textbf{Lab 10: Capstone: File System Simulation} \
Focuses on the integration of all learned concepts, designing a Command-Line File System simulation using OOP principles and Advanced Data Structures.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
