\rhead{HS101: English Communication}
\phantomsection 
\section*{English Communication (HS101)}
\label{course:HS101}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 3 (0-3-0)

\subsection*{Course Content}
\noindent\textbf{Unit 1} \\
\textit{Functional Grammar and Corrections:} Remedial Grammar: Revisiting Tenses (Present, Past, Future) to ensure correct sentence timing. Subject-Verb Agreement drills. Articles (a, an, the) and Prepositions (in, on, at, by). Common Errors in English: Identifying and correcting "Indianisms" and translation errors. Sentence Structure: Simple, Compound, and Complex sentences.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Phonetics and Pronunciation:} The Sounds of English: Introduction to Vowels and Consonants (Basic IPA). Syllables and Word Stress: Understanding where to place emphasis in words (e.g., Pho-TO-gra-phy vs. Pho-to-GRAPH-ic). Intonation and Rhythm: Rising and Falling tones to convey emotion and questions. Neutral Accent training: Reducing Mother Tongue Influence (MTI).

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Listening and Comprehension:} Active Listening Drills: Listening to audio clips (news, simple dialogues) and answering questions. Note-taking: Techniques to capture key points during a conversation. Vocabulary Building: Synonyms, Antonyms, Homophones, and basic idioms. Contextual guessing: Understanding words based on the surrounding sentence.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Basic Speaking and Conversation:} Breaking the Ice: Self-Introductions (Name, Background, Hobbies). Small Talk: Greetings, asking about health/weather, and ending conversations politely. Situational Dialogues: Role-playing everyday scenarios (e.g., Asking for directions, Ordering food, Talking to a teacher). JAM (Just A Minute) Sessions: Speaking for 60 seconds on simple topics to build confidence and fluency.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Reading and Basic Writing:} Reading Skills: Skimming and Scanning simple texts/newspaper headlines for information. Reading aloud for clarity and pause management. Basic Writing: Constructing coherent paragraphs with a Topic Sentence. Informal Letter Writing (to friends/family) to practice flow. Transformation of sentences (Active to Passive voice basics).

\newpage
\rhead{Curriculum Schema}