\rhead{UC1004: Yoga Education}
\phantomsection 
\section*{Yoga Education \& Practices (UC1004)}
\label{course:UC1004}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 2 (0-0-2)

\subsection*{Course Guidelines}
\noindent \textit{This course is practical and seminar-based. Students are expected to practice fundamental Asanas and deliver a comprehensive presentation (PPT) on one of the theoretical themes listed below, supported by scientific evidence or case studies.}

\subsection*{Course Content \& Presentation Themes}
\noindent\textbf{Theme 1: Foundations and Philosophy of Yoga} \\
\textit{Topics for Presentation:} Historical evolution of Yoga (Vedic to Modern). Introduction to Patanjali’s Yoga Sutras. The Eight Limbs of Yoga (Ashtanga): Yama, Niyama, Asana, Pranayama, Pratyahara, Dharana, Dhyana, and Samadhi. Misconceptions about Yoga vs. the reality of holistic wellness.

\vspace{0.3cm}

\noindent\textbf{Theme 2: Physical Culture (Asanas) and Anatomy} \\
\textit{Topics for Presentation:} The science of Surya Namaskar (Sun Salutation) and its physiological effects. Classification of Asanas: Standing, Sitting, Supine, and Prone. Impact of Yoga on the Musculoskeletal system (Flexibility, Core strength, Posture correction). Yoga for preventing ergonomic hazards (e.g., "Tech Neck" and Back pain in computer professionals).

\vspace{0.3cm}

\noindent\textbf{Theme 3: Pranayama and Respiratory Physiology} \\
\textit{Topics for Presentation:} The connection between Breath and the Nervous System. Techniques: Nadi Shodhana (Alternate Nostril Breathing), Kapalbhati (Cleaning Breath), and Bhramari (Humming Bee Breath). Scientific analysis of how Pranayama regulates the Autonomic Nervous System (Sympathetic vs. Parasympathetic) and reduces cortisol levels.

\vspace{0.3cm}

\noindent\textbf{Theme 4: Mental Health and Stress Management} \\
\textit{Topics for Presentation:} Concept of Pratyahara (Withdrawal of senses) in a digital age. Meditation techniques (Dhyana) for focus and concentration. Yoga Nidra (Psychic Sleep) for deep relaxation. Role of Yoga in managing anxiety, depression, and insomnia among students.

\vspace{0.3cm}

\noindent\textbf{Theme 5: Applied Yoga and Lifestyle} \\
\textit{Topics for Presentation:} Yoga for Lifestyle Disorders (Diabetes, Hypertension, Obesity). Corporate Yoga: Techniques for stress management in the workplace. Diet and Nutrition: The concept of Mitahara and Sattvic diet. Integration of Yoga values (Truthfulness, Non-violence) into professional ethics.

\newpage
\rhead{Curriculum Schema}