\rhead{RB-EL2: Deep Learning for Robot Perception}
\phantomsection 
\section*{Deep Learning for Robot Perception (RB-EL2)}
\label{elec:RB_2}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{CNN Architectures for Visual Perception:} Convolutional layers and feature hierarchies, AlexNet to ResNet/VGG, Residual learning and skip connections, Depthwise separable convolutions (MobileNet), Dilated convolutions (DeepLab), Grouped convolutions (ShuffleNet), Attention mechanisms (CBAM, SENet), EfficientNet scaling and compound coefficients.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Object Detection and Segmentation:} Two-stage detectors (R-CNN, Fast/Faster R-CNN), One-stage detectors (YOLO v3-v8, SSD, RetinaNet), Anchor-free methods (CornerNet, FCOS), Instance segmentation (Mask R-CNN), Semantic segmentation (FCN, U-Net, DeepLab series), Panoptic segmentation (Panoptic FPN), Real-time detection trade-offs.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{3D Perception and Depth Estimation:} Monocular depth estimation (MiDaS, Depth Anything), Stereo vision and disparity maps, LiDAR point cloud processing (voxelization, PointNet/PointNet++), Multi-view fusion (MVSNet), 3D object detection (VoxelNet, SECOND, PointRCNN), BEV representations (Lift-Splat-Shoot), Sensor fusion (camera-LiDAR-Radar).

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Visual Odometry and SLAM:} Feature-based VO (ORB-SLAM, DSO), Learning-based VO (DeepVO, VIO), Direct methods vs. indirect, Bundle adjustment optimization, Loop closure detection, Visual-inertial odometry (OKVIS, VINS-Mono), Neural radiance fields (NeRF) for dense reconstruction, Semantic SLAM integration.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Robust Perception and Domain Adaptation:} Sim-to-real transfer (CycleGAN, domain randomization), Uncertainty estimation (Bayesian NNs, Monte Carlo dropout), Test-time adaptation, Continual learning for perception, Multi-modal fusion (BEVFusion, UniFusion), Edge deployment optimization (TensorRT, ONNX Runtime), Anomaly detection in perception pipelines.

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: CNN Implementation from Scratch} \\
Focuses on PyTorch convolutional layers, training on CIFAR/ImageNet subsets.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Object Detection Pipeline} \\
Focuses on YOLOv8/MMDetection, custom dataset annotation/training.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Semantic Segmentation} \\
Focuses on DeepLabv3/HRNet, Cityscapes/KITTI datasets.

\vspace{0.2cm}

\noindent\textbf{Lab 4: 3D Point Cloud Processing} \\
Focuses on Open3D/PointNet, LiDAR segmentation/detection.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Visual Odometry} \\
Focuses on ORB-SLAM3 implementation, KITTI/TUM RGB-D evaluation.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Sensor Fusion} \\
Focuses on nuScenes dataset, camera-LiDAR fusion networks.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Domain Adaptation} \\
Focuses on CycleGAN sim-to-real, BDD100K/Syn2Real experiments.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Real-Time ROS2 Integration} \\
Focuses on perception nodes, NVIDIA Jetson deployment.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Uncertainty Quantification} \\
Focuses on evidential deep learning, out-of-distribution detection.

\vspace{0.2cm}

\noindent\textbf{Lab 10: End-to-End Perception System} \\
Focuses on complete robotic pipeline with custom hardware.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
