\rhead{RB-EL5: Advanced Motion Planning}
\phantomsection 
\section*{Advanced Motion Planning (RB-EL5)}
\label{elec:RB_5}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Geometric Motion Planning:} Configuration space (C-space obstacles), Workspace vs. C-space, Collision detection (GJK algorithm, bounding volume hierarchies), Sampling-based planners (PRM: probabilistic roadmap, lazy evaluation), RRT family (RRT-Connect, RRT*: asymptotic optimality), Informed sampling (elliptical projections).

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Optimization-Based Planning:} Trajectory optimization (CHOMP, STOMP, TrajOpt), Convex optimization (convex MPC, SOCP), Differential flatness for drones/manipulators, Whole-body planning (centroidal dynamics), Contact-rich manipulation (MuJoCo, QP trajectory optimization), Real-time receding horizon control.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Learning for Motion Planning:} Imitation learning (BC, DAGGER, GAIL), Reinforcement learning planners (AlphaPilot, Dreamer), Neural motion policies (MPNet, MotionPolicy), Scene-aware planning (LEGO, SPN), Foundation models for planning (RT-2, Octo), Fast planning via regression (Latent Planners), Sample efficiency techniques.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Multi-Robot and Human-Aware Planning:} Decentralized planning (conflict-based search CBS), Priority planning, Velocity obstacles (ORCA), Human-robot cohabitation (social costs, reciprocal collision avoidance), Dynamic environments (DWA, TEb), Legible and predictable trajectories for HRI.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Advanced Topics:} Nonholonomic planning (dubins paths, Reeds-Shepp curves), Underactuated systems (acrobot, pendubot), Legged locomotion (MPC, WBF), Soft robotics planning, Uncertainty-aware planning (POMDPs, belief space), Integrated perception-planning (VPG, PEPLAN), Real-time certification (iCORA, anytime algorithms).

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: Sampling-Based Planners} \\
Focuses on OMPL implementation (RRT*, PRM), MoveIt! integration.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Optimization Trajectory Generation} \\
Focuses on TOPP-RA, CHOMP in Drake/MuJoCo.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Manipulation Planning} \\
Focuses on Panda arm pick-place, collision-free IK.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Quadrotor Motion Planning} \\
Focuses on minimum-snap trajectories, PX4 SITL.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Multi-Robot Coordination} \\
Focuses on CBS/ORCA in Gazebo swarm.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Neural Planners} \\
Focuses on MPNet fine-tuning, scene-conditioned planning.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Human-Aware Navigation} \\
Focuses on Nav2 social layers, pedestrian datasets.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Legged Robot Locomotion} \\
Focuses on ANYmal MPC, contact sequence optimization.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Real-Time Hardware Demo} \\
Focuses on Jetson deployment, latency optimization.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Complete Planning Pipeline} \\
Focuses on perception-motion-action loop validation.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
