\rhead{RB-EL6: Manipulation \& Grasping Mechanisms}
\phantomsection 
\section*{Manipulation \& Grasping Mechanisms (RB-EL6)}
\label{elec:RB_6}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Rigid Body Kinematics and Statics:} Forward/inverse kinematics for manipulators, Denavit-Hartenberg (DH) parameters, Product of exponentials (POE) formulation, Jacobian matrices (linear/angular velocity), Singularity analysis (manipulability ellipsoid), Static equilibrium (wrench closure), Force/torque relationships.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Grasp Planning and Analysis:} Grasp quality metrics (force closure, form closure), Grasp wrench space (GWS) and manipulability, 2D/3D grasp synthesis (independent contact regions), Friction cone constraints (LMI formulation), Optimal grasp planning (convex optimization), Parallel jaw vs. multifinger grasps, Underactuated hands.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Dexterous Manipulation:} In-hand manipulation (finger gaiting, regrasp), Rolling contact kinematics, Part mating (chamferless assembly), Fixturing and form closure, Hybrid position/force control, Impedance/admittance control, Variable admittance for compliant manipulation.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Contact Dynamics and Control:} Coulomb friction model (stick-slip), Complementarity formulations (LCP, NCP), Time-stepping integrators (MuJoCo, Drake), Hybrid zero dynamics for manipulation, Tactile sensing and feedback, Slip detection and grip adjustment, Soft robotics grasping (continuum models).

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Learning-Based Grasping:} Data-driven grasp prediction (DexNet, GraspNet), 6-DoF grasp detection (deep CNNs, PointNet++), Sim-to-real transfer (domain randomization, system ID), Reinforcement learning for manipulation (QT-Opt, RLBench), Vision-based servoing (IBVS, 6D pose estimation), Model-based planning (MPC with contacts).

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: Manipulator Kinematics} \\
Focuses on DH modeling, Jacobian computation, singularity analysis.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Grasp Quality Metrics} \\
Focuses on GWS computation, force closure testing.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Parallel Jaw Grasping} \\
Focuses on MoveIt! grasp planning, Panda gripper simulation.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Dexterous Hand Simulation} \\
Focuses on Shadow Hand, finger gaiting in Gazebo.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Contact-Rich Manipulation} \\
Focuses on peg insertion, force control tuning.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Vision-Based Grasping} \\
Focuses on DexNet-4, RGB-D grasp detection.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Tactile Sensing Integration} \\
Focuses on DigiTac sensors, slip detection algorithms.

\vspace{0.2cm}

\noindent\textbf{Lab 8: RL Manipulation Policies} \\
Focuses on Stable-Baselines3, custom gym environments.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Real Robot Experiment} \\
Focuses on Franka Panda, pick-and-place validation.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Complete Manipulation Pipeline} \\
Focuses on perception-planning-execution with uncertainty handling.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
