\rhead{MA102: Probability \& Statistics}
\phantomsection 
\section*{Probability \& Statistics (MA102)}
\label{course:MA102}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 3 (3-0-0)

\subsection*{Course Content}
\noindent\textbf{Unit 1} \\
\textit{Foundations of Probability:} Sample spaces, Events, Axiomatic definition of Probability. Permutations and Combinations in complex counting problems. Conditional Probability, Independence of events, and Total Probability Theorem. \textbf{Bayes’ Theorem} and its applications in classification and diagnostic systems (Naïve Bayes foundation). Introduction to Random Variables: Definition, Cumulative Distribution Functions (CDF), and Probability Mass/Density Functions (PMF/PDF).

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Random Variables and Distributions:} Theoretical distributions essential for data modeling. \textit{Discrete:} Bernoulli, Binomial, Geometric, and Poisson (modeling rare events/cyber attacks). \textit{Continuous:} Uniform, Exponential (reliability/waiting times), Gamma, Beta, and **Normal (Gaussian) Distribution** (properties and the Empirical Rule). Expectation, Variance, Skewness, Kurtosis, and Moment Generating Functions (MGF). Transformation of Random Variables.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Multivariate Analysis and Information Theory:} Jointly distributed random variables (Discrete and Continuous). Marginal and Conditional distributions. Independent Random Variables. Covariance, Correlation Coefficient, and Rank Correlation. \textbf{Information Theory for ML:} Self-Information, Entropy (Shannon), Joint and Conditional Entropy, Mutual Information, and **Kullback-Leibler (KL) Divergence** (foundations for Cross-Entropy Loss in Neural Networks).

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Statistical Inference and Estimation:} Sampling distributions: Chi-square, t-distribution, and F-distribution. Central Limit Theorem (CLT) and Law of Large Numbers. \textbf{Parameter Estimation:} Point estimation using **Maximum Likelihood Estimation (MLE)** and **Maximum A Posteriori (MAP)** (Bayesian approach). Interval Estimation: Confidence Intervals for mean, variance, and proportions. Hypothesis Testing: Null/Alternative hypotheses, Type I/II errors, p-values, Z-test, t-test, and ANOVA.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Stochastic Processes for RL and Cyber:} Introduction to Stochastic processes. \textbf{Markov Chains:} State space, Transition Probability Matrix (TPM), Chapman-Kolmogorov equations, Stationary distributions, and Ergodicity (Mathematical foundation of **Reinforcement Learning**). Random Walks and Poisson Processes (Network traffic modeling). Introduction to Monte Carlo Methods and Sampling techniques.

\newpage
\rhead{Curriculum Schema}