\rhead{UC2004: Environment Studies}
\phantomsection 
\section*{Environment Studies (UC2004)}
\label{course:UC2004}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 2 (2-0-0)

\subsection*{Course Content}
\noindent\textbf{Unit 1} \\
\textit{Introduction to Environmental Studies \& Ecosystems:} Foundations: Multidisciplinary nature of environmental studies; Scope and importance; Concept of sustainability and sustainable development. Ecosystem Dynamics: Structure and function of ecosystem; Energy flow in an ecosystem: food chains, food webs and ecological succession. Case Studies: Characteristics of Forest ecosystem, Grassland ecosystem, Desert ecosystem, and Aquatic ecosystems (ponds, streams, lakes, rivers, oceans, estuaries).

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Natural Resources: Renewable and Non-renewable Resources:} Land Resources: Land use change; Land degradation, soil erosion and desertification. Deforestation: Causes and impacts due to mining and dam building on environment, forests, biodiversity and tribal populations. Water Resources: Use and over-exploitation of surface and ground water, floods, droughts, conflicts over water (international \& inter-state). Energy Resources: Renewable and non-renewable energy sources, use of alternate energy sources, growing energy needs, case studies.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Biodiversity and Conservation:} Levels of Diversity: Genetic, species and ecosystem diversity; Biogeographic zones of India; Biodiversity patterns and global biodiversity hot spots. Biodiversity in India: India as a mega-biodiversity nation; Endangered and endemic species of India. Threats and Conservation: Habitat loss, poaching of wildlife, man-wildlife conflicts, biological invasions; In-situ and Ex-situ conservation of biodiversity. Ecosystem Services: Ecological, economic, social, ethical, aesthetic and Informational value.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Environmental Pollution, Policies \& Practices:} Pollution and Waste: Types, causes, effects and controls of Air, water, soil and noise pollution; Nuclear hazards and human health risks; Solid waste management (urban and industrial). Global Issues: Climate change, global warming, ozone layer depletion, acid rain and impacts on human communities and agriculture. Environmental Laws: Environment Protection Act; Air (Prevention \& Control of Pollution) Act; Water Act; Wildlife Protection Act; Forest Conservation Act. Agreements and Rights: International agreements (Montreal and Kyoto protocols, CBD); Nature reserves, tribal populations and rights, and human wildlife conflicts in Indian context.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Human Communities and the Environment:} Population and Society: Human population growth: Impacts on environment, human health and welfare; Resettlement and rehabilitation of project affected persons; case studies. Disaster Management: Floods, earthquake, cyclones and landslides. Environmental Movements: Chipko, Silent valley, Bishnois of Rajasthan. Ethics and Communication: Role of Indian and other religions/cultures in environmental conservation; Environmental communication and public awareness (e.g., CNG vehicles in Delhi).

\newpage
\rhead{Curriculum Schema}