\rhead{DS-EL2: Business Intelligence \& Data Warehousing}
\phantomsection 
\section*{Business Intelligence \& Data Warehousing (DS-EL2)}
\label{elec:DS_2}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Data Warehousing Concepts and Architecture:} Operational Data Store (ODS) vs. data warehouse, Bill Inmon vs. Ralph Kimball approaches (normalized vs. star schema), Data warehouse architecture (staging, ETL, presentation layer), conformed dimensions and fact granularity, Slowly Changing Dimensions (SCD Type 1-6), Surrogate keys and dimensionality modeling.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{OLAP and Multidimensional Analysis:} Online Analytical Processing (OLAP) operations (slice/dice, drill-down/up, pivot), MOLAP vs. ROLAP vs. HOLAP, OLAP cube construction and sparse cube optimization, Bitmap indexing for high-cardinality dimensions, Aggregation hierarchies and roll-up strategies, Materialized views and summary tables.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{BI Tools and Visualization:} Self-service BI platforms (Tableau, Power BI, Looker), Dashboard design principles and storytelling, KPI definition and cascading, Advanced visualizations (small multiples, sparklines, heatmaps), Geographic visualization and custom polygons, Interactive filtering and parameter actions, Embedded analytics and white-labeling.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{ETL/ELT Pipelines and Data Quality:} Extract-Transform-Load vs. Extract-Load-Transform paradigms, Change Data Capture (CDC) techniques (log-based, timestamp, triggers), Data quality dimensions (accuracy, completeness, consistency, timeliness), Data profiling and anomaly detection, Master Data Management (MDM) strategies, Data lineage and impact analysis.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Modern Data Platforms and Lakehouse:} Data lake architectures and schema-on-read challenges, Lakehouse paradigm (Delta Lake, Apache Iceberg, Hudi), Semantic layer abstraction (dbt, LookML), Headless BI and metric stores (Metrics Layer), Real-time analytics (Kafka Streams, Flink SQL), Federated query engines (Presto, Trino, Athena).

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: Dimensional Modeling Workshop} \\
Focuses on star/snowflake schema design, SCD implementation, surrogate key generation.

\vspace{0.2cm}

\noindent\textbf{Lab 2: OLAP Cube Development} \\
Focuses on multidimensional cube construction, aggregation hierarchies, slice/dice operations.

\vspace{0.2cm}

\noindent\textbf{Lab 3: BI Dashboard Creation} \\
Focuses on Tableau/Power BI storytelling, KPI dashboards, interactive parameters/filters.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Advanced Visualizations} \\
Focuses on custom polygons, small multiples, geographic heatmaps, trend analysis.

\vspace{0.2cm}

\noindent\textbf{Lab 5: ETL Pipeline Implementation} \\
Focuses on Apache Airflow DAGs, CDC with Debezium, data transformation patterns.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Data Quality Framework} \\
Focuses on Great Expectations validation, anomaly detection, data profiling.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Lakehouse Architecture Setup} \\
Focuses on Delta Lake tables, time travel, ACID transactions, schema evolution.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Semantic Layer Development} \\
Focuses on dbt transformations, LookML models, metric definitions.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Real-Time Analytics Pipeline} \\
Focuses on Kafka + Flink streaming analytics and materialized views.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Complete BI Solution} \\
Focuses on end-to-end data warehouse to executive dashboard delivery.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
