\rhead{DS-EL3: Mining Massive Datasets}
\phantomsection 
\section*{Mining Massive Datasets (DS-EL3)}
\label{elec:DS_3}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Scalable Data Mining Frameworks:} MapReduce programming model revisited, Beyond MapReduce (Spark, Dask), Data stream mining challenges (concept drift, memory constraints), Sliding window models, Reservoir sampling, Massive data partitioning strategies, Approximate query processing and sampling guarantees, Sketching algorithms fundamentals.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Frequent Itemset Mining and Association Rules:} Apriori algorithm and candidate generation, FP-growth and FP-tree structure, Eclat and vertical data format, Sampling-based frequent itemset mining, Parallel/distributed FP-growth, Association rule interestingness measures (support, confidence, lift, conviction), Sequential pattern mining (GSP, SPADE, PrefixSpan).

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Graph Mining and Network Analysis:} Graph representation (adjacency lists, CSR format), PageRank algorithm and variants (weighted, personalized), HITS algorithm, Triangle counting and clustering coefficients, Community detection (Louvain, spectral clustering, label propagation), Graph sampling and streaming algorithms, Subgraph isomorphism and motif discovery.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Dimensionality Reduction and Clustering at Scale:} Mini-batch k-means and scalable EM, Canopy clustering, BIRCH hierarchical clustering, Spectral clustering approximation, t-SNE/UMAP for large datasets, Random projection trees, Streaming PCA and incremental SVD, CUR matrix decomposition for interpretability.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Near-Duplicates and Recommendation Systems:} MinHash and Locality Sensitive Hashing (LSH) for Jaccard similarity, SimHash for text documents, Near-duplicate detection at web scale, Matrix factorization (SVD, ALS, NTF), Neighborhood-based collaborative filtering, Content-based recommendation, Hybrid recommenders, Scalable bandit algorithms (LinUCB, Thompson sampling).

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: Scalable MapReduce/Spark Implementation} \\
Focuses on distributed word count, log analysis, and custom partitioners.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Frequent Itemset Mining} \\
Focuses on Apriori/FP-growth comparison, parallel implementation, association rules.

\vspace{0.2cm}

\noindent\textbf{Lab 3: PageRank and Graph Analytics} \\
Focuses on GraphFrames/GraphX implementation, triangle counting, community detection.

\vspace{0.2cm}

\noindent\textbf{Lab 4: LSH for Near-Duplicates} \\
Focuses on MinHash/LSH indexing, recall-precision tradeoff analysis.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Streaming Data Processing} \\
Focuses on Spark Streaming, sliding window aggregations, anomaly detection.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Scalable Clustering} \\
Focuses on mini-batch k-means, BIRCH, spectral clustering approximation.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Matrix Factorization Recommenders} \\
Focuses on ALS implementation, cold-start handling, evaluation metrics.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Graph Sampling and Streaming} \\
Focuses on reservoir sampling, random walk sampling, streaming PageRank.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Approximate Query Processing} \\
Focuses on sampling guarantees, error bounds, progressive sampling.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Massive Dataset Mining Pipeline} \\
Focuses on complete scalable mining solution from ingestion to insights.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
