\rhead{SE-EL2: Agile Project Governance \& Management}
\phantomsection 
\section*{Agile Project Governance \& Management (SE-EL2)}
\label{elec:SE_2}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-1-0)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Agile Governance Frameworks and Principles:} Agile Manifesto values and principles revisited, Governance vs. control in agile contexts, SAFe (Scaled Agile Framework) levels (team, program, portfolio), LeSS (Large Scale Scrum), Nexus framework, Spotify engineering model, Disciplined Agile Delivery (DAD), Comparative analysis of scaling frameworks, Governance maturity models for agile adoption.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Program and Portfolio Management:} Program Increment (PI) planning in SAFe, ART (Agile Release Train) synchronization, Weighted Shortest Job First (WSJF) prioritization, OKR (Objectives and Key Results) alignment, Portfolio Kanban for value stream management, Lean portfolio management, Investment funding and guardrails, Strategic themes and portfolio epics.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Enterprise Agile Metrics and Analytics:} Flow metrics (cycle time, lead time, throughput), Value stream mapping and bottleneck analysis, Cumulative Flow Diagrams (CFD), Predictability metrics (planned vs. actual velocity), Quality metrics (defect escape rate, DORA metrics), Business value delivered vs. planned, OKR achievement tracking, Data-driven decision making dashboards.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Risk Management and Compliance in Agile:} Agile risk management (product, technical, organizational risks), Risk-adjusted planning and probabilistic forecasting, Monte Carlo simulation for sprint capacity, Compliance by design (regulatory guardrails, audit trails), Governance ceremonies (PI planning, System Demos, Inspect \& Adapt), Decision rights and escalation paths, Architectural runway governance.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Organizational Transformation and Change:} Agile assessment models (Shu-Ha-Ri, Satir Change Model), Enterprise agile coaching frameworks, Transformation backlogs and roadmaps, Communities of Practice (CoP) establishment, Leadership alignment and executive sponsorship, Metrics for transformation success, Scaling anti-patterns and failure modes, Sustainable agile adoption strategies.

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: SAFe PI Planning Simulation} \\
Focuses on Program Increment planning, dependency mapping, and risk assessment workshop.

\vspace{0.2cm}

\noindent\textbf{Lab 2: WSJF Prioritization Framework} \\
Focuses on business value/cost/time criticality analysis for portfolio backlog prioritization.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Flow Metrics Dashboard Development} \\
Focuses on implementing cycle time, throughput, and CFD visualizations using real project data.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Monte Carlo Sprint Forecasting} \\
Focuses on probabilistic velocity prediction and confidence intervals for release planning.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Value Stream Mapping Workshop} \\
Focuses on identifying bottlenecks and waste in software delivery pipelines.

\vspace{0.2cm}

\noindent\textbf{Lab 6: OKR Cascade Implementation} \\
Focuses on organizational goal alignment from portfolio to team level.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Agile Maturity Assessment} \\
Focuses on organizational agile assessment and transformation roadmap development.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Risk-Adjusted Backlog Management} \\
Focuses on incorporating uncertainty and risk into sprint and release planning.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Governance Dashboard Creation} \\
Focuses on executive reporting with key agile metrics and portfolio health indicators.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Enterprise Agile Simulation} \\
Focuses on full-scale agile transformation simulation with multiple teams and dependencies.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
