\rhead{SE-EL5: Enterprise Application Architecture}
\phantomsection 
\section*{Enterprise Application Architecture (SE-EL5)}
\label{elec:SE_5}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-1-0)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Enterprise Architecture Frameworks and TOGAF:} The Open Group Architecture Framework (TOGAF) ADM cycle, Architecture Development Method phases (preliminary, architecture vision, business/data/application/technology architecture), Architecture repository and content framework, Enterprise continuum and solutions landscape, Architecture governance and compliance, Zachman framework comparison, TOGAF certification levels.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Domain-Driven Design for Enterprise Applications:} Strategic DDD patterns (bounded contexts, context maps, core/ supporting/ generic domains), Ubiquitous language establishment, Aggregate design principles (consistency boundaries, transactional consistency), Domain events and event storming workshops, Anti-corruption layer patterns, Hexagonal architecture and ports/adapters, CQRS and Event Sourcing implementation.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Enterprise Integration Patterns:} Messaging patterns (publish-subscribe, point-to-point, request-reply), Enterprise Service Bus (ESB) vs. modern alternatives, API management lifecycle (design, security, versioning, deprecation), Canonical data models and contract-first design, Transaction strategies (Sagas, choreography, orchestration), Data consistency patterns (eventual consistency, 2PC alternatives).

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Microservices and SOA Governance:} Service inventory and capability mapping, Service contracts and SLAs, Service versioning strategies (major/minor/semver), Circuit breaker and bulkhead patterns, Service mesh governance (Istio, Consul), Centralized logging and distributed tracing, Golden signals monitoring for enterprise services, Multi-tenancy architecture patterns.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Enterprise Reference Architectures:} Java EE reference architecture (EJB, JPA, JSF, CDI), Spring Boot enterprise patterns, .NET Core enterprise architecture, Cloud-native enterprise patterns (12-factor apps, strangler pattern), Legacy modernization strategies (strangler fig, branch by abstraction), Reference architectures (J2EE Petstore, Netflix OSS stack), Architecture decision records (ADRs).

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: TOGAF ADM Workshop} \\
Focuses on Architecture Vision creation, stakeholder map development, and governance framework design.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Domain-Driven Design Event Storming} \\
Focuses on bounded context identification, aggregate modeling, and ubiquitous language workshops.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Enterprise Integration Patterns} \\
Focuses on messaging patterns implementation with Apache Camel/Kafka Connect.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Hexagonal Architecture Implementation} \\
Focuses on ports/adapters pattern with domain-driven core and multiple technology adapters.

\vspace{0.2cm}

\noindent\textbf{Lab 5: CQRS/ES Implementation} \\
Focuses on event sourcing, materialized views, and command/query separation.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Service Mesh Enterprise Deployment} \\
Focuses on Istio multi-cluster setup, traffic management, and observability.

\vspace{0.2cm}

\noindent\textbf{Lab 7: API Management Platform} \\
Focuses on API gateway configuration, versioning, rate limiting, and developer portal.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Multi-Tenancy Architecture} \\
Focuses on database-per-tenant vs. shared database strategies implementation.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Legacy Modernization Patterns} \\
Focuses on strangler pattern implementation and branch-by-abstraction techniques.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Enterprise Reference Architecture} \\
Focuses on complete enterprise application following TOGAF/DDD patterns with production deployment.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
