\rhead{CX301: Computer Networks Lab}
\phantomsection 
\section*{Computer Networks Lab (CX301)}
\label{course:CX301}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 2 (0-0-2)

\subsection*{Course Content}
\noindent\textbf{Lab 1: Network Device Study and Cable Crimping} \\
Focuses on identifying network devices (hub, switch, router), crimping straight-through and crossover cables, verifying connectivity using cable testers.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Basic Network Commands and Troubleshooting} \\
Focuses on using ping, traceroute, nslookup, netstat, ifconfig/ipconfig, analyzing connectivity issues and DNS resolution problems.

\vspace{0.2cm}

\noindent\textbf{Lab 3: IPv4 Subnetting and IP Addressing} \\
Focuses on calculating subnets, configuring IP addresses/subnets/VLSM, verifying connectivity across different subnets using Packet Tracer.

\vspace{0.2cm}

\noindent\textbf{Lab 4: ARP Protocol Analysis} \\
Focuses on capturing ARP requests/replies using Wireshark, configuring static ARP entries, analyzing ARP cache poisoning scenarios.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Ethernet Frame Analysis} \\
Focuses on capturing Ethernet frames with Wireshark, analyzing frame structure (preamble, header, CRC), observing CSMA/CD behavior.

\vspace{0.2cm}

\noindent\textbf{Lab 6: VLAN Configuration and Trunking} \\
Focuses on configuring VLANs on managed switches, setting up trunk links (802.1Q), inter-VLAN routing using router-on-a-stick.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Static Routing Configuration} \\
Focuses on configuring static routes between multiple routers, verifying routing tables, analyzing path determination using show ip route.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Dynamic Routing Protocols (RIP/OSPF)} \\
Focuses on configuring RIP and OSPF, observing route convergence, analyzing routing protocol metrics and neighbor relationships.

\vspace{0.2cm}

\noindent\textbf{Lab 9: TCP 3-Way Handshake and Connection States} \\
Focuses on capturing TCP handshake packets with Wireshark, configuring socket programs to observe SYN/ACK/FIN states.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Network Simulation and Congestion Analysis} \\
Focuses on building multi-router topologies in Packet Tracer, generating traffic to observe congestion, analyzing throughput vs. packet loss.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
