\rhead{CX304: Database Management Systems Lab}
\phantomsection 
\section*{Database Management Systems Lab (CX304)}
\label{course:CX304}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 2 (0-0-2)

\subsection*{Course Content}
\noindent\textbf{Lab 1: DDL and Basic DML Operations} \\
Focuses on creating databases, tables with constraints (PRIMARY KEY, FOREIGN KEY, CHECK, UNIQUE), inserting, updating, and deleting records with basic SELECT queries.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Advanced SQL Queries} \\
Focuses on nested queries, multiple JOIN types (INNER, LEFT, RIGHT, FULL OUTER), aggregate functions with GROUP BY/HAVING clauses.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Subqueries and Set Operations} \\
Focuses on correlated subqueries, EXISTS/IN/NOT IN operators, UNION/INTERSECT/MINUS operations, and view creation/manipulation.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Stored Procedures and Functions} \\
Focuses on creating PL/SQL procedures, functions, triggers, and packages for business logic encapsulation.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Transaction Control and Concurrency} \\
Focuses on implementing ACID transactions, SAVEPOINT usage, demonstrating lost update and dirty read anomalies.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Isolation Levels Testing} \\
Focuses on configuring different transaction isolation levels and observing phantom reads, non-repeatable reads behavior.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Index Analysis and Performance} \\
Focuses on creating various index types (B-tree, bitmap), analyzing execution plans, measuring query performance improvements.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Query Optimization Study} \\
Focuses on rewriting queries for better performance, using EXPLAIN PLAN, creating materialized views.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Database Backup and Recovery} \\
Focuses on full/incremental backups, point-in-time recovery, log analysis, and disaster recovery scenarios.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Complete Application Database Design} \\
Focuses on end-to-end database design (ER diagram → normalized schema → complex queries → stored procedures → performance tuning).

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
