\rhead{ML-EL2: MLOps \& Production AI}
\phantomsection 
\section*{MLOps \& Production AI (ML-EL2)}
\label{elec:ML_2}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{MLOps Fundamentals and ML Lifecycle:} ML project lifecycle stages (data, training, validation, deployment, monitoring), MLOps maturity levels (manual, ML automation, continuous ML), Model drift types (concept drift, data drift, upstream drift), Golden ML pipelines, Experiment tracking and reproducibility, Versioning strategies (data, models, code, environment), ML metadata stores and lineage tracking.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Data Management and Feature Engineering:} Data versioning (DVC, Delta Lake), Data quality gates and schema validation, Feature stores (Feast, Tecton, Hopsworks), Online/offline feature serving, Feature drift detection, Data lineage and impact analysis, Automated data validation (Great Expectations, Deequ), PII/PHI data anonymization and compliance.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Model Development and Experimentation:} MLflow tracking server and experiment management, Weights \& Biases (W\&B) sweeps and hyperparameter optimization, Ray Tune distributed tuning, Model registries and staging environments, A/B testing frameworks (Optimizely, Flagr), Canary deployments and shadow testing, Model performance baselines and champion/challenger patterns.

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Model Deployment and Serving:} Model packaging (ONNX, TorchServe, TensorFlow Serving), Containerization best practices (Docker multi-stage builds), Kubernetes inference servers (KFServing, Seldon Core, KServe), Serverless ML (AWS SageMaker Serverless, Cloud Run), Edge deployment (TensorFlow Lite, ONNX Runtime), Multi-model serving and GPU sharing, Batch vs. real-time inference.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{ML Monitoring, Governance, and Continuous Training:} Model monitoring (KServe model monitoring, Prometheus integration), Drift detection algorithms (KS2 test, PSI, MMD), Automated retraining triggers and CI/CD for ML, Human-in-the-loop validation, Model explainability (SHAP, LIME integration), Bias/fairness monitoring (AIF360, Fairlearn), Compliance frameworks (GDPR, HIPAA ML controls), Blue-green ML deployments.

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: ML Pipeline with MLflow} \\
Focuses on complete ML lifecycle tracking, experiment comparison, and model registry usage.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Feature Store Implementation} \\
Focuses on Feast feature store setup, online/offline serving, and feature drift monitoring.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Automated Hyperparameter Tuning} \\
Focuses on Ray Tune sweeps, parallel training, and optimization visualization.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Model Deployment with KServe} \\
Focuses on Kubernetes inference serving, traffic splitting, and autoscaling.

\vspace{0.2cm}

\noindent\textbf{Lab 5: A/B Testing Infrastructure} \\
Focuses on canary deployments, statistical significance testing, and rollout analysis.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Model Monitoring and Drift Detection} \\
Focuses on drift detection pipelines, alerting configuration, and performance degradation analysis.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Continuous Training Pipeline} \\
Focuses on automated retraining triggers, model validation, and champion/challenger selection.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Explainable AI Integration} \\
Focuses on SHAP/LIME integration with production models and drift-aware explanations.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Multi-Model Serving Platform} \\
Focuses on GPU sharing, dynamic model loading, and inference cost optimization.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Production ML Platform} \\
Focuses on complete MLOps platform with data versioning, automated training, deployment, and monitoring.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
