\rhead{ML-EL6: Probabilistic Graphical Models}
\phantomsection 
\section*{Probabilistic Graphical Models (ML-EL6)}
\label{elec:ML_6}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-1-0)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Probabilistic Graphical Models Fundamentals:} Directed (Bayesian networks) vs. undirected (Markov Random Fields) models, Factorization of joint distributions, Conditional independence and d-separation, Markov properties and Hammersley-Clifford theorem, Graphical model representation advantages, Exponential family distributions and sufficiency, Common probability distributions in graphical models.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Inference Algorithms:} Exact inference (variable elimination, belief propagation), Complexity analysis of inference, Approximate inference (importance sampling, likelihood weighting), Markov Chain Monte Carlo (MCMC) methods (Metropolis-Hastings, Gibbs sampling), Sequential Monte Carlo (particle filters), Loopy belief propagation and mean field approximation.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Learning Graphical Models:} Parameter learning in Bayesian networks (maximum likelihood, MAP estimation), Dirichlet priors and Dirichlet-multinomial models, Structure learning (score-based, constraint-based), K2 algorithm and greedy hill-climbing, Markov blanket discovery, PC algorithm for DAG structure learning, Learning Markov Random Fields (pseudolikelihood estimation).

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Dynamic Bayesian Networks and Hidden Markov Models:} Hidden Markov Models (HMM) - three fundamental problems (evaluation, decoding, learning), Forward-backward algorithm, Viterbi algorithm, Baum-Welch EM algorithm, Dynamic Bayesian Networks (DBN) for time series, Switching Kalman filters, Continuous-time graphical models.

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Advanced Topics and Applications:} Gaussian graphical models and precision matrix estimation, Conditional Random Fields (CRF) for structured prediction, Factor graphs and sum-product algorithm, Nonparametric Bayesian models (Indian Buffet Process), Deep generative models with graphical structure, Causal discovery from observational data, Applications in bioinformatics, NLP, and robotics.

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: Bayesian Network Inference} \\
Focuses on exact inference using variable elimination and belief propagation.

\vspace{0.2cm}

\noindent\textbf{Lab 2: MCMC Sampling Methods} \\
Focuses on Metropolis-Hastings, Gibbs sampling convergence diagnostics.

\vspace{0.2cm}

\noindent\textbf{Lab 3: Hidden Markov Model Implementation} \\
Focuses on Viterbi path decoding, forward-backward probabilities, EM learning.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Parameter Learning in BNs} \\
Focuses on MLE/MAP estimation with missing data and Dirichlet priors.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Structure Learning Algorithms} \\
Focuses on score-based and constraint-based DAG learning from data.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Dynamic Bayesian Networks} \\
Focuses on time series modeling and filtering/smoothing applications.

\vspace{0.2cm}

\noindent\textbf{Lab 7: Conditional Random Fields} \\
Focuses on sequence labeling (NER, POS tagging) with CRF models.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Gaussian Graphical Models} \\
Focuses on precision matrix estimation and partial correlation networks.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Factor Graph Inference} \\
Focuses on sum-product algorithm for complex graphical structures.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Graphical Models Application} \\
Focuses on complete PGM solution for domain-specific prediction task.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
