\rhead{ML-EL1: Reinforcement Learning}
\phantomsection 
\section*{Reinforcement Learning (ML-EL1)}
\label{elec:ML_1}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 4 (3-0-2)

\subsection*{Theory}
\noindent\textbf{Unit 1} \\
\textit{Reinforcement Learning Foundations:} Markov Decision Processes (MDPs) formalization (states, actions, transition probabilities, rewards), Bellman equations and optimality principle, Model-free vs. model-based RL, Exploration vs. exploitation tradeoff ($\epsilon$-greedy, UCB, entropy regularization), Discounted future rewards and value iteration, Policy iteration algorithm, Generalized Policy Iteration (GPI).

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Model-Free Temporal Difference Learning:} Monte Carlo methods (first-visit, every-visit, importance sampling), Temporal Difference (TD) learning (TD($\theta$), TD($\lambda$)), Q-Learning algorithm and off-policy learning, SARSA and on-policy control, Double Q-Learning and expected SARSA, N-step bootstrapping methods, Eligibility traces implementation.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Function Approximation and Deep RL:} Linear function approximation (average reward, gradient TD), Tile coding and radial basis functions, Deep Q-Networks (DQN) architecture (experience replay, target networks), Double DQN improvements, Dueling DQN and distributional RL, Rainbow DQN combining multiple improvements, Policy gradient methods (REINFORCE, actor-critic).

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Policy Gradient Methods and Actor-Critic:} REINFORCE algorithm with baseline subtraction, Advantage Actor-Critic (A2C/A3C), Proximal Policy Optimization (PPO) - clipped objective, Trust Region Policy Optimization (TRPO), Soft Actor-Critic (SAC), Deterministic Policy Gradients (DPG), Twin Delayed DDPG (TD3), Continuous control benchmarks (MuJoCo, PyBullet).

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Advanced RL Topics and Applications:} Multi-agent RL (MARL) cooperation/competition, Centralized training with decentralized execution (CTDE), Self-play and population-based training, Model-based RL (Dyna, MBPO, Dreamer), Hierarchical RL (options framework, Feudal Networks), Offline RL (conservative Q-learning, behavior regularization), RLHF (RL from Human Feedback) for LLM alignment.

\newpage

\subsection*{Practical}
\noindent\textbf{Lab 1: MDPs and Dynamic Programming} \\
Focuses on value iteration, policy iteration, and MDP environment implementation.

\vspace{0.2cm}

\noindent\textbf{Lab 2: Monte Carlo and TD Learning} \\
Focuses on MC/TD methods comparison on classic environments (Cliff Walking, Taxi).

\vspace{0.2cm}

\noindent\textbf{Lab 3: Q-Learning and SARSA Implementation} \\
Focuses on off-policy/on-policy control, exploration strategies, and convergence analysis.

\vspace{0.2cm}

\noindent\textbf{Lab 4: Deep Q-Network (DQN) Training} \\
Focuses on Atari games, experience replay, target networks, and hyperparameter tuning.

\vspace{0.2cm}

\noindent\textbf{Lab 5: Policy Gradient Methods} \\
Focuses on REINFORCE implementation for discrete/continuous action spaces.

\vspace{0.2cm}

\noindent\textbf{Lab 6: Actor-Critic Algorithms} \\
Focuses on A2C/A3C parallel training and advantage estimation techniques.

\vspace{0.2cm}

\noindent\textbf{Lab 7: PPO and TRPO Implementation} \\
Focuses on constrained policy optimization and sample efficiency comparison.

\vspace{0.2cm}

\noindent\textbf{Lab 8: Continuous Control with SAC/TD3} \\
Focuses on MuJoCo robotics environments and state-of-the-art continuous RL.

\vspace{0.2cm}

\noindent\textbf{Lab 9: Multi-Agent RL Scenarios} \\
Focuses on cooperative/competitive MARL with self-play training.

\vspace{0.2cm}

\noindent\textbf{Lab 10: Custom RL Environment Project} \\
Focuses on complete RL agent development for robotics/navigation application.

\vspace{0.2cm}

\newpage
\rhead{Curriculum Schema}
