\rhead{MA201: Discrete Math}
\phantomsection 
\section*{Discrete Mathematical Structures (MA201)}
\label{course:MA201}

\noindent \small{\hyperref[main:scheme]{$\leftarrow$ Back to Scheme}} \vspace{0.5cm}

\noindent \textbf{Credits:} 3 (3-0-0)

\subsection*{Course Content}
\noindent\textbf{Unit 1} \\
\textit{Mathematical Logic and Proofs:} Propositional Logic: Statements, Connectives, Truth Tables, Tautologies, and Contradictions. Normal Forms (CNF, DNF). Predicate Logic: Quantifiers ($\forall, \exists$), Nested Quantifiers, and Inference Theory. Proof Techniques: Direct, Indirect (Contrapositive), Contradiction, and \textbf{Mathematical Induction} (Strong and Weak) with applications to algorithm correctness.

\vspace{0.3cm}

\noindent\textbf{Unit 2} \\
\textit{Sets, Relations, and Functions:} Set Theory: Operations, Power Sets, Principle of Inclusion-Exclusion. \textbf{Relations:} Properties (Reflexive, Symmetric, Transitive), Equivalence Relations and Partitions (Database foundations). Partial Order Relations: Posets, Hasse Diagrams, Lattices, and Chains. Functions: Injective, Surjective, Bijective, Inverse Functions, and the Pigeonhole Principle with applications to hashing.

\vspace{0.3cm}

\noindent\textbf{Unit 3} \\
\textit{Combinatorics and Recurrence Relations:} Counting Principles: Permutations, Combinations, and Derangements. Binomial and Multinomial Theorems. \textbf{Recurrence Relations:} Generating Functions, Linear Recurrence Relations with Constant Coefficients (Homogeneous and Non-homogeneous). Solving recurrences for \textbf{Algorithm Analysis} (e.g., Fibonacci, Divide and Conquer complexity).

\vspace{0.3cm}

\noindent\textbf{Unit 4} \\
\textit{Graph Theory:} Fundamentals: Directed and Undirected graphs, Weighted graphs, Paths, Cycles, and Connectivity. Eulerian and Hamiltonian Paths (Network routing problems). Graph Representation (Adjacency Matrix/List). Planar Graphs and Graph Coloring (Map coloring, Register allocation). Isomorphism and Homomorphism. Shortest Path algorithms concepts (Dijkstra/Floyd-Warshall context).

\vspace{0.3cm}

\noindent\textbf{Unit 5} \\
\textit{Trees and Algebraic Structures:} \textbf{Trees:} Rooted Trees, Binary Search Trees, Spanning Trees (Prim’s and Kruskal’s logic), and Tree Traversals. \textbf{Algebraic Structures:} Groups, Subgroups, Cyclic Groups, Lagrange’s Theorem. Rings and Fields (Finite Fields). Introduction to Number Theory: Modular Arithmetic, Euclidean Algorithm, Fermat’s Little Theorem, and Euler’s Totient Function (Mathematical foundation for \textbf{RSA Cryptography}).

\newpage
\rhead{Curriculum Schema}